Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-03-13 inst/surfing_jsa_rebuttal.tex

Referee 1

To the authors,

An interesting paper submitted, thank you. My only
real concerns are your actual comments related to the sport of
competitive surfing.

Firstly, some confusion as you initially state heats involve one or
two other competitiors. However, you begin the second paragraph in the
introduction with a statement saying 'up to four surfers', so which is
it?

Second statement is that the World Surf League is not the main
governing body of surfing competition. What about the International
Surfing Association, who organised and ran the Olympics, also run the
World Surfing games and national level competitions. The WSL really
only run professional level competition.

34 surfers are for the mens competition.

Why are you ignoring the "wildcards". No real justification for
this. Especially as some of the wildcards competed in more than one of
the competitions. So some form of statement to validate why you omit
the wildcards would be good.

One other comment is based around the statements around judging in the
second last paragraph before section 1.1. Judging is based on the
heats conditions, as surfing is performed within an ever changing
environment. Therefore, to try and compare between surfers throughout
a competition is really not applicable. As each heat, and each wave is
different from the previous. So a direct comparison throughout a
competition is somewhat unfair for the competitors. As for the claim
around abilities, is this not subjective, as each performer is judged
upon their surfing prowess on that particular wave in that particular
heat.

Your hypothesis on Brazilians preference for reef breaking waves is based on what?

Statement on the points system not affecting competitiors
behaviour. How could you say this, as depending on the finishing spot
at each competition, the athlete is awarded a set number of points. So
a higher placing in a competition will award you a greater number of
points. Impacting upon your placing in the rankings for the subsequent
competition. And therefore, your seeding for the next competition.

Other than that, an interesting paper. Just feel you need to clarify a
few of the points above to provide readers a clearer viewpoint of the
sport and understanding to the importance of your study. As I found it
quite interesting. Just some claims and anomalies related to
competitive surfing.

Referee 2

The World Surf League (WSL) is the main governing
body of surfing competitions (World Surf League 2021).

Not exactly. What about ISA (world surfing games, Olympics, national
championships, regional championships).

The WSL conducts a
world tour in which the best ranked 34 surfers compete;

This only constitutes for male competition

in addition, at each tour venue, two "wildcard" surfers also enter the
competition, who are ignored here

Why are they ignored? Especially as some of the wildcards competed in multiple events.

Up to four surfers--the competitors--are in the water simultaneously,
watched by up to five judges.

You stated one or two other competitors above. So which is it?

However, for a particular heat, the order statistic--that is, which
competitor scored most highly, second highest, and third--is
informative about competitors abilities: the wave environment is
common to each surfer

But judging is based on the heats conditions, as surfing is performed
within an ever changing environment. Therefore, to try and compare
between surfers throughout a competition is really not applicable. As
each heat, and each wave is different from the previous. So a direct
comparison is somewhat unfair for the competitors.

and here we apply it to assess the supposed Brazilian preference for
reef-breaking waves.

Based on what hypothesis?

the points system itself is intrinsically arbitrary: the details of
the points system does not affect the competitors' behaviour but can
change the overall ranking;

How could you say this, as depending on finishing spot at each
competition, the athlete is awarded a set number of points. So a
higher placing in a competition will award you a greater number of
points. Impacting upon your placing in the rankings for the subsequent
competition. And therefore, your seeding for the next competition.

### 2022-03-17 inst/surfing_jsa_rebuttal.tex

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

\section*{MS596 ``Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods" by Driver and Hankin: rebuttal}

Below, the reviewer's comments are in black, and our replies to the
issues are in \textcolor{blue}{blue}.  We have indicated changes to
the manuscript where appropriate.

Short story: we have accommodated all the comments with rewording and
clarification.  The resulting document is, I believe, stronger and
more scholarly than before and I recommend it to you.

\section*{Detailed rebuttal: Reviewer \#1}

@@

Second statement is that the World Surf League is not the main
governing body of surfing competition. What about the International
Surfing Association, who organised and ran the Olympics, also run the
World Surfing games and national level competitions. The WSL really
only run professional level competition.

{\bf intent: }  WSL is the *professional* body organisation.

34 surfers are for the mens competition.

{\bf intent: } We focus on the male competition for the moment

@@

One other comment is based around the statements around judging in the
second last paragraph before section 1.1. Judging is based on the
heats conditions, as surfing is performed within an ever changing
environment. Therefore, to try and compare between surfers throughout
a competition is really not applicable. As each heat, and each wave is
different from the previous. So a direct comparison throughout a
competition is somewhat unfair for the competitors. As for the claim
around abilities, is this not subjective, as each performer is judged
upon their surfing prowess on that particular wave in that particular
heat.

{\bf intent:}  Will put some statistical comment in here.

Your hypothesis on Brazilians preference for reef breaking waves is based on what?

{\bf intent: } We tested for Brazilians NOT liking reef breaking waves
and liking point and beach breaking waves.  consistent with 6.1

### 2022-03-18 inst/surfing_jsa_rebuttal.tex

and here we apply it to assess the supposed Brazilian preference for
reef-breaking waves.

Based on what hypothesis?

the points system itself is intrinsically arbitrary: the details of
the points system does not affect the competitors' behaviour but can
change the overall ranking;

How could you say this, as depending on finishing spot at each
competition, the athlete is awarded a set number of points. So a
higher placing in a competition will award you a greater number of
points. Impacting upon your placing in the rankings for the subsequent
competition. And therefore, your seeding for the next competition.

\begin{verbatim}
             nation wct01 wct02 wct03 wct04 wct05 wct06 wct07 wct08 wct09 wct10 wct11 points
 Ferreira       BRA     1     5    17     5    17     2    17     9     2     1     1  59740
 Medina         BRA     5     5    17    17     5     1     2     1     9     9     2  56475
 Smith          ZAF     3     3    17     5     2     9     3     5     9     2    17  49985
 Toledo         BRA     9     2     5    17     1     3     9     2    17     5    17  49145
 Andino         USA     2    17     5     2     3     3    17    17     5     5     9  46655
 Igarashi       JPN     9     9     1     9     5     5    17     9    17     3    17  40185
 Florence       HAW     3     1    17     1     5   INJ   INJ   INJ   INJ   INJ     5  37700
 Slater         USA    33     5     3     9     9     9    17     9    17     9     3  34845
 Wright         AUS     9     9    17     9    17     5     1     3    17    17    17  34780
 Flores         FRA    17     9     2    17    33    17     5    17     1    17    33  32515
 Wilson         AUS    17    17     9     3     5    17     9     5     9    17     9  31515
 Moniz          HAW     5     9    17     5    17    17     3    17     9    33     9  29525
 Bourez         FRA     9    17     9     9     9     9     9    33     9    17     5  29315
 Callinan       AUS    17     3     9     5    17     9    17    17     5    33    17  27535
 Freestone      AUS    17    17     9    17    17    17     9    17     3     5     5  27535
 Colapinto      USA    17   INJ    17    17     9    17     9     3    17     9     3  27450
 Ibelli         BRA    33    17    17     3    17    17     5    17    17     3     9  26885
 Carmichael     AUS     5    17     5    33     9    17    17     9     9     9    17  26760
 Buchan         AUS    17    17     5    33    33     5    17     5     5    17    17  25630
 Coffin         USA     5     9     9     9    17    17    17    17    17     9    17  23345
 Crisanto       BRA    17     9    17     9    33     9    17    33    17     5     9  23345
 Dora           BRA     9    17    17    17    17    17    17     5     9    17     5  22780
 Silva          BRA    17     9    17    17     9     9     9     9    33    17    17  21920
 Cardoso        BRA     9     9    17    17    17     9    17     9    17    17    17  19930
 Mendes         BRA    17    33     9    17     9    33    17    17    17     9     9  19930
 Rodrigues      BRA    17    17     3    17     9    17    33    33    17     9    33  19640
 Zietz          HAW    17   PAR    33     9    17     5    17     9    33    17    17  18300
 Duru           FRA    17    33     9    17     9    17     9    17    17    17    17  17940
 Lau            HAW    17    17    33    17    17     9    17     9     9    33    17  17940
 Bailey         AUS    17    17    33    17    17    17    17    17    17     9     9  15950
 Fioravanti     ITA    33    17     9    17   INJ   INJ   INJ   INJ     3    17    33  14455
 Andre          BRA    33    17    17    17    17    33     5    17    33    17    17  14320
 Christie       NZL    17    17    17    17    17    17    17    17    33    33     9  13960
 Morais         PRT     -     -     -    33     3    17    33     -    17    17    33  10870
 deSouza        BRA   INJ   INJ   INJ   INJ    17    17     5   INJ   INJ   INJ   INJ   8995
 Wright_M       AUS     9    17    17   INJ   INJ   INJ   INJ   INJ   INJ   INJ   INJ   7570
 Willcox        AUS     -     5    33    33     -     -     -     -     -     -     -   5275
 Lacomare       FRA     -     -     -     -     -     -     -     -     5     -     -   4745
 Heazlewood     AUS     9    17     -     -     -     -     -     -     -     -     -   4650
 Robinson       AUS     -     -     -     9     -     -     -     -     -     -     -   3320
 Vaast          FRA     -     -     -     -     -     -     9     -     -     -     -   3320
 Colapinto_C    USA     -     -     -     -     -     -     -    17     -    17     -   2660
 Herdy          BRA    17     -     -     -   INJ     -     -    33     -     -     -   1860
 Couzinet       FRA     -     -     -     -     -    33     -     -    17     -     -   1595
 Waida          INA     -     -    17     -     -     -     -     -     -     -     -   1330
 Kymerson       BRA     -     -     -     -    17     -     -     -     -     -     -   1330
 February       ZAF     -     -     -     -     -    17     -     -     -     -     -   1330
 Mamiya         HAW     -     -     -     -     -     -     -    17     -     -     -   1330
 Matson         USA     -     -     -     -     -     -     -    17     -     -     -   1330
 Schilling      USA     -     -     -     -     -     -     -    17     -     -     -   1330
 Mignot         FRA     -     -     -     -     -     -     -     -    17     -     -   1330
 Ribeiro_V      PRT     -     -     -     -     -     -     -     -     -    17     -   1330
 Blanco         PRT     -     -     -     -     -     -     -     -     -    17     -   1330
 deVault        HAW     -     -     -     -     -     -     -     -     -     -    17   1330
 Kemper         HAW     -     -     -     -     -     -     -     -     -     -    17   1330
 Mann           AUS     -    33     -     -     -     -     -     -     -     -     -    265
 Huxtable       AUS     -    33     -     -     -     -     -     -     -     -     -    265
 Ribeiro_A      BRA     -     -     -     -    33     -     -     -     -     -     -    265
 DeVries        ZAF     -     -     -     -     -    33     -     -     -     -     -    265
 Newton         HAW     -     -     -     -     -     -    33     -     -     -     -    265
 Drollet        PYF     -     -     -     -     -     -    33     -     -     -     -    265
\end{verbatim}
\end{document}

### 2022-03-19 inst/surfing_jsa_rebuttal.tex

{\bf intent: } We focus on the male competition for the moment

\subsection*{Ignoring wildcards}

Why are you ignoring the "wildcards". No real justification for
this. Especially as some of the wildcards competed in more than one of
the competitions. So some form of statement to validate why you omit
the wildcards would be good.

\textcolor{blue}{See the table of results at the end of this document.
  We see that the wildcards participate in very few venues: of the 25
  wildcards, 20 appear at only a single venue, four at two venues, and
  only one (Willcox) at three venues.  Further, the wildcards perform
  poorly: of the 31 ranked scores, only 5 are better than last or
  second to last.  Maximum likelihood Bradley-Terry strengths for such
  competitors are either exactly zero or very small (here $\sim
  10^{-5}$) and are neither interesting from a surfing lore
  perspective, nor susceptible to meaningful statistical
  inference.\newline Quite apart from the fact that our focus was on
  the top-performing competitors, including such datapoints is
  problematic for several reasons.  Firstly, numerical optimization
  techniques tend to struggle as one increases the number of players:
  one is necessarily working with a higher-dimensional problem,
  leading to poorly defined maxima [the Jacobian becomes very small];
  in addition, local maxima become more numerous and difficult to
  identify.\newline We have added a brief summary of this to the
  manuscript.}

### 2022-03-19 inst/surfing.Rmd
referee: why ignore wildcards?

>>>>

Some quick analysis of the wildcard surfers.  First, how did the wildcards perform?

```{r}
wilds <- as.matrix(surfing_table[-(1:36),2:12])
wilds[wilds == "INJ"] <- NA
wilds[wilds == "-"] <- NA
table(wilds)
```

Above we see that 10 wildcards placed "33" [meaning last place], 16
placed "17" [second to last] three placed "9" and two placed "5".
[the placings are treated as character strings which is why they are
not in numerical orer].  Next, how many venues did each wildcard
participate at?

```{r}
table(apply(wilds,1,function(x){sum(!is.na(x))}))
```

We see that 20 wildcards appeared at 1 venue, 4 at two venues, and one
(Willcox) at three venues.q

### 2022-03-19 inst/surfing_jsa.tex
referee: why ignore wildcards?

Surfing is a popular and growing activity in terms of both
participation and viewing \citep{warshaw2010}.  Competitive surfing
involves surfers competing against one or two other surfers in heats
lasting 30-50 minutes \citep{booth1995}.  Points are awarded on a
10-point scale: each of five judges awards points to a surfer riding a
particular wave.  The World Surf League (WSL) is the main governing
body of surfing competitions \citep{wsl}.  The WSL conducts a world tour in
which the best ranked 34 surfers compete; in addition, at each tour
venue, two ``wildcard" surfers also enter the competition, who are
ignored here.  The 2019 WSL tour saw competitive events held at eleven
different surf locations.

>>>>

Surfing is a popular and growing activity in terms of both
participation and viewing \citep{warshaw2010}.  Competitive surfing
involves surfers competing against one or two other surfers in heats
lasting 30-50 minutes \citep{booth1995}.  Points are awarded on a
10-point scale: each of five judges awards points to a surfer riding a
particular wave.  The World Surf League (WSL) is the main governing
body of surfing competitions \citep{wsl}.  The WSL conducts a world
tour in which the best ranked 34 surfers compete; in addition, at each
tour venue, two ``wildcard" surfers also enter the competition, who
are ignored here due to their generally poor performance and low
participation.

The 2019 WSL tour saw competitive events held at eleven
different surf locations.
